---
title: "diversity model fit"
output: html_document
date: "2024-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r source_scripts}
source("./diversityModelFunctions.R")
```

## Define priors

We use a grid design with 100 communities (as in the original model paper Suzuki et al) to explore the parameter space. We define the priors for the parameters of the model. 

We fix the number of species and communities:
```{r}
n_com<- 100 #number of communities (same as Suzuki et al.)
n_sp<-127 #as per Boag et al. 2016
```

Prior ranges
```{r}
nu_range <- c(0, 0.1)
aid_range <- c(5,6,7)
wval_range <- c(0.1, 2)
hval_range <- c(0.01, 10)
```

## Sample from the prior

```{r}
set.seed(123) # for repro
n_sim <- 10000
sim_res <- list(nu = runif(n_sim, nu_range[1], nu_range[2]),
                   aid = sample(aid_range, n_sim, replace = TRUE),
                   wval = runif(n_sim, wval_range[1], wval_range[2]),
                   hval = exp(runif(n_sim, log(hval_range[1]), (hval_range[2]))),
                alpha_sims = matrix(NA, nrow =100, ncol = n_sim))
```

```{r}
for (i in 1:n_sim) {
  sim_res$alpha_sims[, i] <- changing.values.a(nu = sim_res$nu[i], aid = sim_res$aid[i], wval = sim_res$wval[i], hval = sim_res$hval[i])
}
# create date in format for file name (including time)
date=format(Sys.time(), "%Y%m%d%H%M") # removed time 
saveRDS(sim_res, paste0("simulation_results", date, ".rds"))
```


## Compare simulations with observed
```{r}
sim_res <- readRDS("simulation_results202411201401.rds")


```

```{r}
alpha_observed <- c(20, 36, 28)
sim_res$alpha_sims[sim_res$alpha_sims == 0] <- NA
alpha_sim_mean <- colMeans(sim_res$alpha_sims, na.rm = TRUE)
```

Count simulations where everything went extinct:
```{r}
sum(is.na(alpha_sim_mean))
```

Plot the distribution of the mean alpha diversity from the simulations
```{r}
hist(alpha_sim_mean)
```
```{r}
library(abc)
sim_param<- data.frame(nu=sim_res$nu, aid=sim_res$aid, wval=sim_res$wval, hval=sim_res$hval)
avalon_rej <- abc::abc(target = 20,param = sim_param,sumstat = alpha_sim_mean, tol=0.5, method="neuralnet")
```
```{r}
dimnames(avalon_rej$unadj.values)[2]
```

```{r}
hist(avalon_rej$unadj.values[,4])

```

```{r}
white_rej <- abc::abc(target = 36,param = sim_param,sumstat = alpha_sim_mean, tol=0.5, method="neuralnet")
```
```{r}
nama_rej <- abc::abc(target = 28,param = sim_param,sumstat = alpha_sim_mean, tol=0.5, method="neuralnet")


```

Rejection distributions for the three periods
```{r, fig.width=4, fig.height=20}}
par(mfrow=c(3,1))
i <- 3
hist(avalon_rej$unadj.values[,i], breaks=20)
hist(white_rej$unadj.values[,i], breaks=20)
hist(nama_rej$unadj.values[,i], breaks=20)

```

```{r, fig.width=4, fig.height=20}}
par(mfrow=c(3,4))
i <- 3
hist(avalon_rej, breaks=20, ask=FALSE)
hist(white_rej, breaks=20, ask=FALSE)
hist(nama_rej, breaks=20, ask=FALSE)

```

Summary of the posteriors
```{r}
summary(avalon_rej)
summary(white_rej)
summary(nama_rej)


```

Now plot the posteriors as densities:
```{r}
#head(avalon_rej$adj.values)
par(mfrow=c(3,1))
plot(density(avalon_rej$adj.values[,3]), main="Avalon", xlab="alpha", ylab="Density")
plot(density(white_rej$adj.values[,3]), main="White Sea", xlab="alpha", ylab="Density")
plot(density(nama_rej$adj.values[,3]), main="Nama", xlab="alpha", ylab="Density")

```

